---
title: "Comparison of Cylindar Count to Engine Displacement"
author: "Isla Fisher"
date: "31/07/2025"
output: pdf_document
---

A client named Big Dawg has commissioned this work with the research question: 
Do cars with more cylinders not have bigger engines?

The goal of this report is to provide Big Dawg with clarity as to if cars with 
more cylinders have bigger or not.

This report will investigate the relationship between engine displacement and 
number of cylinders by using linear regression and the tools available on 
R studio.


```{=latex}
\vspace{2cm}
```


## Creating a Cylinder and Engine Displacement Table

```{r} 
# Calling the tidyverse package.
library(tidyverse)
```


```{r}
# Calling mtcars and assigning it to the object mtcars.
mtcars <- mtcars
```


```{r}
# Creating Table 1 and assigning it to the object my_table. This involves 
# selecting the columns cylinder and displacement from mtcars data set. Then a 
# new column is created for Table 1 with the name disp_m representing 
# displacement mean.

my_table <- mtcars |>
  select(cyl, disp) |>
  summarise(disp_m = round(mean(disp), 2), 
            .by=cyl)
```


```{r}
# Formatting my_table to match the lab instructions. Changing the names of the 
# columns in Table 1 and using kable to style it correctly.
my_table <- my_table |>
  rename(Cylinders = cyl, 
         Displacement = disp_m)


knitr::kable(my_table, align = 'c', caption = 
               "Mean engine displacement by number of cylinders") |>
  kableExtra::kable_styling(full_width = FALSE, position = "center")

```


```{=latex}
\vspace{2cm}
```

## Creating a Cylinder and Engine Displacement Bar Chart

```{r}
# Creating and formatting the Bar Chart to align with the lab instructions.
my_table |>
  ggplot(aes(x = Cylinders, y = Displacement)) +
  geom_bar(stat = "identity") +
  labs(title = "Figure 1. Mean engine displacement by number of cylinders",
       x = "Number of Cylinders", 
       y = "Engine Displacement (cm^3)") +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),              # Remove full panel border
    axis.line = element_line(color = "black")
  )
```

Figure 1 shows that as cylinders increase, engine displacement also increases. This shows that Big Dawgs original statement is false and cars with more cylinders do have bigger engines.